Clouds and Sensitivities Across a Hierarchy of GFDL CMIP6 Models

### 2020-11-18 Silvers_cfmip.tex
modify abstract for the U Wy seminar

\begin{abstract}
Observations of cloud fraction from multiple satellites are compared to the clouds simulated by 
GFDL's new AM4 climate model.  
The new atmospheric climate model succeeded at significantly reducing the TOA radiative flux biases as compared 
to CERES observations.  Despite a relatively low top-of-atmosphere sensitivity to uniform warming of SSTs 
(Cess-type warming experiments), the corresponding coupled climate model, CM4, has high transient and 
equilibrium climate sensitivities.

A systematic picture is presented of the modeled clouds across a hierarchy of model configurations which utilize 
the AM4 and CM4 climate models.  This hierarchy includes the CFMIP Aquaplanet and \textit{amip} experiments, 
fully coupled model experiments as well as additional \textit{amip}-like experiments with particular SST patterns.  
We demonstrate the large range of sensitivities that are possible from a single parent atmospheric climate model.   
Looking at the global mean radiative feedbacks across the different model configurations as well as in the context 
of CMIP5 and CMIP6 models will allow us to assess to what extent the cloud feedbacks in the idealized 
experiments relate to the fully coupled experiments and to observed clouds.  
\end{abstract}

@@

\begin{table}
\begin{center}
\caption{Global mean radiative feedbacks ($\rm{W\, m^{-2} K^{-1}}$).  Values in the first four columns are calculated as differences for 
particular years (see text).  The values for the \textit{abrupt 4x$\rm{CO}_2^{b}$} and  \textit{abrupt 4x$\rm{CO}_2^{c}$} experiments
have been calculated using linear regression of 1 year means over years 1-150 and 51-300, respectively.  
\textit{abrupt 4x$\rm{CO}_2^{a}$}, \textit{abrupt 4x$\rm{CO}_2^{b}$}, and \textit{abrupt 4x$\rm{CO}_2^{c}$} represent data from the same experiment, the feedback values have simply been calculated differently.
}
    \begin{tabular}{*{7}{c}}
    \hline
    \hline
 Feedback     &  \textit{amip }  & \textit{historical}   &  \textit{1pctCO$_2$} & \textit{abrupt-4xCO$_2^{a}$} & \textit{abrupt-4xCO$_2^{b}$} & \textit{abrupt-4xCO$_2^{c}$}  \\ \hline
    Net $\alpha$    & -1.77      &   -1.83      &  -1.22          &  -1.11    &  -0.84   &   -0.55              \\ 
    \\
    Net CRE  &    -0.46            &   -0.45       &  -0.16          & -0.06   & 0.17      &   0.62         \\  
    \\
    LW CRE    &   -0.03           &   -0.14        &  0.10            & 0.14   &  0.25        & 0.11          \\  
    \\
    SW CRE  &     -0.43             &    -0.31       &  -0.26            & -0.20    &  -0.08      &  0.51         \\  
    \\
    LW CLR   &     -2.25             &    -1.94       &  -1.71            & -1.70    &   -1.79      & -1.74        \\  
    \\
    SW CLR  &      0.91            &     0.56        & 0.65              & 0.65    &    0.79         & 0.57          \\  
    \\
    $\Delta \rm{T}_s$     &     1.3     & 0.8        &  4.3           & 5.7     &   3.90 (7.8?)        &   4.60 (9.2?)         \\  \hline

>>>>

\begin{abstract}

A systematic picture will be presented of the simulated clouds across a hierarchy of CMIP6 model configurations  that have been run with 
the AM4 and CM4 climate models from GFDL.  This hierarchy is composed of both CMIP6 and CFMIP (Cloud Feedback Model Intercomparison Project) experiments that include an ocean covered planet, atmosphere/land only Earth-like experiments, and fully coupled model experiments both of the historical era and future warming scenarios (\textit{aqua-control, amip, historical, amip-p4K, amip-future4K, abrupt-4xCO2}, and \textit{1pctCO2}).   
We demonstrate the large range of climate sensitivities that can result from a single parent atmospheric climate model.   
Despite a relatively low top-of-atmosphere sensitivity to uniform warming of SSTs 
(Cess-type warming experiments), the corresponding coupled climate model, CM4, has high transient (2 K) and 
equilibrium climate sensitivities (~ 5K).  Cloud feedbacks play a large role in setting these sensitivities.  
Looking at the global mean radiative feedbacks across the different model configurations as well as in the context 
of CMIP5 and CMIP6 models will allow us to assess to what extent the cloud feedbacks in the idealized 
experiments relate to the fully coupled experiments and to the clouds observed in Earth's atmosphere.

Observations of cloud fraction from multiple satellite products (CALIPSO, MISR, MODIS, and ISCCP)  are compared to the clouds simulated by the AM4 and CM4 climate models.  
The AM4 and CM4 climate models succeeded at significantly reducing the TOA radiative flux biases as compared 
to CERES-EBAF observations.  Partly as a result of this realistic representation of the TOA energetics the coupled model
improved the representation of precipitation, reduced the Intertropical Convergence Zone bias, and produces an excellent El Nino-Southern Oscillation.  However, we find large biases in cloud fraction relative to observations across wide regions of the globe and throughout the depth of the atmosphere.

I will discuss some of the challenges that are inherent to improving cloud simulations in GCMs.
This work points to the need for drawing dependable cloud constraints from the observations we have.  The important challenge will then be to implement the cloud constraints while continuing to maintain a realistic representation of the radiative fluxes at the TOA of global models.

\end{abstract}

@@

\begin{table}
\begin{center}
\caption{Global mean radiative feedbacks ($\rm{W\, m^{-2} K^{-1}}$).  Values in the first four columns are calculated as differences for 
particular years (see text).  The values for the \textit{abrupt 4x$\rm{CO}_2^{b}$} and  \textit{abrupt 4x$\rm{CO}_2^{c}$} experiments
have been calculated using linear regression of 1 year means over years 1-150 and 51-300, respectively.  
\textit{abrupt 4x$\rm{CO}_2^{a}$}, \textit{abrupt 4x$\rm{CO}_2^{b}$}, and \textit{abrupt 4x$\rm{CO}_2^{c}$} represent data from the same experiment, the feedback values have simply been calculated differently.
}
    \begin{tabular}{*{7}{c}}
    \hline
    \hline
 Feedback     &  \textit{amip }  & \textit{historical}   &  \textit{1pctCO$_2$} & \textit{abrupt-4xCO$_2^{a}$} & \textit{abrupt-4xCO$_2^{b}$} & \textit{abrupt-4xCO$_2^{c}$}  \\ \hline
    Net $\alpha$    & -1.77      &   -1.83      &  -1.22          &  -1.11    &  -0.84   &   -0.55              \\ 
    \\
    Net CRE  &    -0.46            &   -0.45       &  -0.16          & -0.06   & 0.17      &   0.62         \\  
    \\
    LW CRE    &   -0.03           &   -0.14        &  0.10            & 0.14   &  0.25        & 0.11          \\  
    \\
    SW CRE  &     -0.43             &    -0.31       &  -0.26            & -0.20    &  -0.08      &  0.51         \\  
    \\
    LW CLR   &     -2.25             &    -1.94       &  -1.71            & -1.70    &   -1.79      & -1.74        \\  
    \\
    SW CLR  &      0.91            &     0.56        & 0.65              & 0.65    &    0.79         & 0.57          \\  
    \\
    $\Delta \rm{T}_s$     &     1.3     & 0.8        &  4.3           & 5.7     &   3.90 (7.8)        &   4.60 (9.2)         \\  \hline
